---
title: "Wrangling Question: How to Add a Totals Row?"
author: "Dominguez Center for Data Science Workshop"
date: "3-21-2025"
format: pdf
execute:
  warning: false
  message: false
editor_options: 
  chunk_output_type: console
---


# Load Packages



```{r}
library(tidyverse)
```


# Load the Data

Run the following code to load the data.

```{r}
# Load the data
colleges <- read_csv("data/ccbasic21.csv") 

```


# Question: How do I add a line at the bottom of a table that includes the values for all observations?

* For each of the Mid-Atlantic states, let's produce a table that contains the average admission rate and number of schools.  
* Then let's add a row with the average admissions rate for all of the schools in the Mid-Atlantic.

```{r}
by_school <- colleges %>%
  filter(STABBR %in% c("PA", "NJ", "VA", "MD", "DE", "WV", "DC")) %>%
  drop_na(ADM_RATE) %>%
  group_by(STABBR) %>%
  summarize(avg_admit = mean(ADM_RATE), count = n()) %>%
  arrange(avg_admit)

mid_atlantic <- colleges %>%
  filter(STABBR %in% c("PA", "NJ", "VA", "MD", "DE", "WV", "DC")) %>%
  drop_na(ADM_RATE) %>%
  summarize(avg_admit = mean(ADM_RATE), count = n()) %>%
  mutate(STABBR = "All") %>%
  relocate(STABBR)

combined_table <- by_school %>%
  bind_rows(mid_atlantic)
```

# Extras

## Question: How do I export as a csv so that I can open it in Excel (or Google Sheets) and make it prettier?

Run the following code:

```{r}
write_csv(combined_table, "data/avg_admit_table.csv")
```


Then within the Files pane in the lower right:

* Click on the "data" folder.
* Check the box next to "avg_admit_table.csv"
* Go to More and select "Export".  This will also you to download the file to your computer.

## Question: Can I make the table look nicer within `R`?

Yes!  Here's an example.

```{r}
library(kableExtra)
kable(combined_table,
             format = "latex",
             booktabs = TRUE, longtable = TRUE, 
             digits = 2,
      col.names = c("State", "Average Admissions Rate", "Number of Schools"))

```

```{r}
library(gt)
combined_table %>%
  gt() %>%
  data_color(columns = avg_admit, method = "numeric",
             palette = "viridis", reverse = TRUE) %>%
  tab_header(title = "Average Admission Rates for Mid-Atlantic Schools",
             subtitle = "Data from College Scorecard") %>%
  tab_source_note(
    source_note = "Focused on schools with a Basic Carnegie Classification of 'Confers majority Baccalaureate degrees with arts and sciences focus'"
  ) %>%
  fmt_number(columns = avg_admit, decimals = 2) %>%
  cols_label(STABBR = md("**State**"), avg_admit = md("**Average Admission Rate**"),
             count = md("**Number of Schools**"))
```

